---
title: "WPBR Model"
author: "Stephen Huysman"
date: "`r Sys.Date()`"
fontsize: 10pt 
output:
  html_document:
    toc: true
    toc_float: true
---

Files:
Parks:
./SITE_LOCATIONS.csv : Working copy of study sites, with all new data added in
./SITE_LOCATIONS_ORIGINAL_FROM_EJ.csv : Original Data from EJ
./SITE_LOCATIONS_ORIGINAL_FROM_EJ_FIX_NAMES.csv : Original Data from EJ with "fivepks" transect names fixed - added 0 before some of the transect numbers to match new numbers received from NPS

Tree Data:
./Complete_Site_Monitoring_Dataset.csv : working copy with all new data added in
./Complete_Site_Monitoring_Dataset_original_from_ej.csv : original from EJ
./Complete_Site_Monitoring_Dataset_original_from_ej_add_dbh.csv : original from EJ with California transect_id's fixed, to use as input for cleaning script
./Complete_Site_Monitoring_Dataset_original_from_ej_add_dbh_clean.csv : file above with NA dbh's for california transects added in

Do spatial random effects account for transect?
https://github.com/pbs-assess/sdmTMB/discussions/192
>My data was collected across multiple plots which are distributed across two forest complexes themselves. This means I have repeated measurements per plot and the plots are nested into a specific forest complex. With glmmTMB you can add this nested random effect as follows: (1|site/plot). Using the same syntax does not work for sdmTMB. Hence, my question: is it possible to add nested random effects to a sdmTMB model? And if so, how do you implement it?

>>You should be able to do the same thing by coding your plot IDs in your
data as unique across all sites not just within sites and then having a
random intercept for site and a random intercept for plot.

>>As a warning, if your sites are spatially arranged and you’re also
including a spatial random field, you might struggle to estimate the model.
One is assuming the sites are IID normally distributed and the other is
assuming they are random fields (MVN with some covariance function).
Theoretically you could have random site noise on top of the random field,
but unless you have a lot of data that may be hard to estimate and you
might want to pick one or the other.


# Load Data
```{r}
library(daymetr)
library(tidyverse)
library(sf)
library(terra)
library(modelr)
library(elevatr)
library(janitor)
library(MuMIn)
library(GGally)
library(lme4)
library(ggeffects)
library(glue)
library(slider)
library(viridis)

crs_wgs84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

parks <- read_csv("./SITE_LOCATIONS.csv")
## parks <- read_csv("./SITE_LOCATIONS_ORIGINAL_FROM_EJ_FIX_NAMES.csv")

##gridmet_data_dir <- "./out/elev_corrected"
gridmet_data_dir <- "./out/"

gridmet_files <- fs::dir_ls(path = gridmet_data_dir, glob = "*.csv")

gridmet_data <- vroom::vroom(gridmet_files, id = "transect_id", altrep = TRUE)

gridmet_data <- gridmet_data %>%
    select(-"...1") %>%
    filter(
        year(Date) >= 2000,
        year(Date) <= 2020
    ) %>%
    mutate(transect_id = str_extract(transect_id, regex(".*/(.*)\\.csv"), group = 1)) %>%
    filter(transect_id != "Butterfield") ## Issue pulling gridmet data for site

## Add rolling mean of RH
gridmet_data <- gridmet_data %>%
    arrange(Date) %>%
    group_by(transect_id) %>%
    mutate(rollRH = slide_dbl(RH, mean, .before = 3, .after = 3))

## Annual Climate
gridmet_data_annual <- gridmet_data %>%
    filter(month(Date) == 8 | month(Date) == 9) %>%
    group_by(transect_id, year(Date)) %>%
    summarize(
        T = mean(T),
        ##T_cor = mean(T_cor),
        meanRH = mean(RH),
        ## sumRH = sum(RH),
        meanVPD = mean(VPD),
        ##sumVPD = sum(VPD),
        maxVPD = max(VPD),
        minVPD = min(VPD),
        P = sum(P),
        rollRH_max = max(rollRH),
        rollRH_min = min(rollRH)
    )

## A-S Climate
gridmet_data_as <- gridmet_data_annual %>% 
    group_by(transect_id) %>%
    summarize(
        asT = mean(T),
        ##asT_cor = mean(T_cor),
        asRH = mean(meanRH),
        ## sumRH = mean(sumRH),
        asVPDmean = mean(meanVPD), ## Mean annual asVPD sum
        asVPDmax = mean(maxVPD),
        asVPDmin = mean(minVPD),
        asP = mean(P), ## Mean annual P sum
        rollRH_max = mean(rollRH_max)
    )


monitoring_data <- read_csv("./Complete_Site_Monitoring_Dataset.csv") %>%
##monitoring_data <- read_csv("./Complete_Site_Monitoring_Dataset_original_from_ej_add_dbh_clean.csv") %>%
    left_join(parks, by = c("network", "transect_id")) %>%
    filter(tree_species == "PIAL" | tree_species == "PINALB") %>%
    filter(transect_id != "Butterfield") %>% ## Issue pulling gridmet data for site
    filter(!is.na(dbh_cm), dbh_cm != 999, dbh_cm != -999, dbh_cm != Inf, dbh_cm != -Inf) %>%
    filter(dbh_cm > 0) %>% ## Not sure what 0 cm dbh means, about 189 trees dbh=0
    filter(!is.na(br_status))


mod_data_tree_level <- left_join(monitoring_data, gridmet_data_as, by = "transect_id")
    ## filter(!is.na(asT), !is.na(asRH), !is.na(asVPD)) ## filter sites with missing climate data

    ### Remove duplicate tree_id in transect.  Keep the first one found.
mod_data_tree_level <- mod_data_tree_level %>%
        filter(year <= 2020 | is.na(year)) %>% ### All glac network observation years are NA
##        filter(year >= 2003 & year <= 2020) %>%
##        filter(tree_status == "L") %>%
        arrange(desc(year)) %>% ### Keep transects from most recent year
        distinct(across(c(transect_id, tree_id)),
            .keep_all = TRUE
        ) %>%
        mutate(
            br_status = factor(br_status)
        )

mod_data_tree_level %>%
    filter(is.na(year)) %>%
    distinct(network, park) ## should only be GLAC

mod_data_tree_level <- mod_data_tree_level %>%
##    left_join(gridmet_data_annual, by = "transect_id") %>%
    mutate(
        transect_id = factor(transect_id),
        park = factor(park),
        network = factor(network),
        log_dbh_cm = log(dbh_cm)
    )


### Convert lat longs to equidistant projection for sdmTMB spatial models
max_lat <- round(max(mod_data_tree_level$lat), 0)
min_lat <- round(min(mod_data_tree_level$lat), 0)
mid_lat <- round(mean(mod_data_tree_level$lat), 0)
mid_long <- round(mean(mod_data_tree_level$long), 0)

crs_lcc <- glue("+proj=lcc +lon_0={mid_long} +lat_1={min_lat} +lat_2={max_lat} +units=km")

mod_data_tree_level <- st_as_sf(mod_data_tree_level, coords = c("long", "lat"), crs = crs_wgs84)

mod_data_tree_level <- st_transform(mod_data_tree_level, crs = crs_lcc)

mod_data_tree_level <- bind_cols(mod_data_tree_level, st_coordinates(mod_data_tree_level)) %>% as_tibble()

```


# Map

```{r map}
## extent <- ext(parks_sf)
library(AOI)
library(maptiles)
library(tidyterra)

## NA_aoi <- aoi_get(x = "North America")

## extent <- ext(NA_aoi)

## base_map <- get_stadiamap(bbox = c(left = -125,
##                                     bottom = 35,
##                                     right = -105,
##                                     top = 55),
##           maptype = "stamen_terrain", 
##           crop = TRUE,
##           zoom = 5) 


wbp_existing <- rast("/home/steve/OneDrive/whitebark/data/existing/wbp_existing.tif")

mod_data_sf <- st_as_sf(mod_data_tree_level, crs = crs(crs_lcc)) %>% st_transform(crs("EPSG:3857"))

base_map <- get_tiles(mod_data_sf, provider = "Esri.WorldTopoMap", crop = TRUE, zoom = 5)

## ggplot() +
##     geom_polygon(data = map_data("state"), aes(long, lat, group = group), color = "black", fill = "white", inherit.aes = FALSE) +
##     geom_polygon(data = map_data("world", "Canada"), aes(long, lat, group = group), color = "black", fill = "white", inherit.aes = FALSE) +
##     geom_spatraster(data = wbp_existing, mapping = aes(fill = wbp_existing), na.rm = TRUE, alpha = 0.3) +
##     scale_fill_continuous(name = "WBP Range", na.value = NA) +
##     geom_sf(data = mod_data_sf, aes(color = park), inherit.aes = FALSE) +
##     labs(color = "Parks") +
##     coord_sf(xlim = c(-125, -105), ylim = c(35, 55), crs = 4326) +
##     ggtitle("Whitebark WPBR Monitoring Sites") +
##     theme_bw(base_size = 16) +
##     scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "W")) +
##     scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N")) +
##     ggspatial::annotation_scale(
##                    location = "tr",
##                    bar_cols = c("grey60", "white"),
##                    text_family = "ArcherPro Book"
##                ) +
##     ggspatial::annotation_north_arrow(
##                    location = "tr", which_north = "true",
##                    pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
##                    style = ggspatial::north_arrow_nautical(
##                                           fill = c("grey40", "white"),
##                                           line_col = "grey20",
##                                           text_family = "ArcherPro Book"
##                                       )
##                )
##ggsave(file.path(ccrp_presentation_dir, "wpbr_map.png"), width = 8, height = 8, type = "cairo")

ggplot() +
    geom_spatraster_rgb(data = base_map) +
    geom_sf(data = mod_data_sf, aes(color = park), inherit.aes = FALSE) +
    ggtitle("WPBR Probability of Infection, 33cm DBH, CanESM2 rcp45 2071-2090") +
    labs(fill = "P(infection)") +
    scale_x_continuous(labels = function(x) paste0(x)) +
    scale_y_continuous(labels = function(x) paste0(x))


ggmap(base_map) +
    geom_sf(data = mod_data_sf, inherit.aes = FALSE, aes(color = park)) +
    coord_sf(crs=4326, expand = FALSE)

```

# A priori models
| Variable | Description               |
|----------|---------------------------|
| asT      | Aug/Sep Mean Temperature  |
| asRH     | Aug/Sep Mean RH           |
| asVPD    | Aug/Sep Sum of VPD        |
| asP      | Aug/Sep Sum of P          |
| dbh_cm   | Diameter at breast height |

- Location variables?

# EDA

```{r}
cols = c("br_status", "asT", "asVPD", "asRH", "rollRH_max", "asP", "dbh_cm", "log_dbh_cm")

ggpairs(mod_data_tree_level, columns = cols)

ggpairs(mod_data_tree_level, mapping = aes(color = park), columns = cols)

mod_data_tree_level %>%
    pivot_longer(cols = c(asVPD, asT, asRH, asP)) %>%
    filter(network != "fivepks") %>%
    ggplot(aes(x = value, y = br_status)) +
    geom_boxplot() +
    facet_wrap(~name, scales = "free")

mod_data_tree_level %>%
    pivot_longer(cols = c(asVPD, asT, asRH, asP)) %>%
    ggplot(aes(x = value, y = br_status)) +
    geom_boxplot() +
    facet_wrap(~name, scales = "free")

```

## Season Plot
Has Aug/September patters of climate shifted?

```{r}
monthly_climate <- gridmet_data %>%
    left_join(parks, by = "transect_id") %>%
    mutate(
        year = year(Date),
        month = month(Date)
    ) %>%
    group_by(year, month, park, transect_id) %>%
    summarize(
        P = sum(P),
        VPD = sum(VPD),
        RH = RH,
        T = T
    ) %>%
    group_by(year, month, park) %>%
    summarize(
        P = mean(P),
        T = mean(T),
        VPD = mean(VPD),
        RH = mean(RH),
        transect_id = transect_id
    )

ggplot(monthly_climate, mapping = aes(x = month, y = RH, color = year, group = year)) +
    geom_line() +
    scale_x_continuous(breaks = 1:12) +
    scale_color_viridis() +
    geom_vline(xintercept = 8) +
    geom_vline(xintercept = 10) +
    facet_wrap(~park)

ggplot(monthly_climate, mapping = aes(x = month, y = VPD, color = year, group = year)) +
    geom_line() +
    scale_x_continuous(breaks = 1:12) +
    scale_color_viridis() +
    geom_vline(xintercept = 8) +
    geom_vline(xintercept = 10) +
    facet_wrap(~park)

ggplot(monthly_climate, mapping = aes(x = month, y = T, color = year, group = year)) +
    geom_line() +
    scale_x_continuous(breaks = 1:12) +
    scale_color_viridis() +
    geom_vline(xintercept = 8) +
    geom_vline(xintercept = 10) +
    facet_wrap(~park)

ggplot(monthly_climate, mapping = aes(x = month, y = P, color = year, group = year)) +
    geom_line() +
    scale_x_continuous(breaks = 1:12) +
    scale_color_viridis() +
    geom_vline(xintercept = 8) +
    geom_vline(xintercept = 10) +
    facet_wrap(~park, scales = "free")
```

# Model Selection

## Test for Covariance
```{r}
glmer_all <- glmer(br_status ~ asT + asVPD + asRH + asP + dbh_cm + (1 | park) + (1 | park:transect_id),
                   data = mod_data_tree_level,
                   family = binomial
)

summary(glmer_all)

cor_table <- vcov(glmer_all) %>%
    cov2cor() %>%
    as.matrix()

cor_table %>%
    knitr::kable()

abs(cor_table) > 0.7

library(corrplot)
corrplot(cor(select(mod_data_tree_level, asT, asVPD, asP, asRH, dbh_cm)), method = "number", type = "upper")

corrplot(cor(select(mod_data_tree_level, max_T, min_T, max_P, min_P, max_VPD, min_VPD, max_RH, min_RH, dbh_cm)), method = "number", type = "upper")

```

Collinear variables (glmer_all):
asT, asVPD
asVPD, asRH

Corrplot:
asT, asVPD
asRH, asP
asVPD, asRH? = -0.57

## Only FE

``` r
glm_t <- glm(br_status ~ asT + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t)

glm_t_2 <- glm(br_status ~ poly(asT, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2)

## Square T?
anova(glm_t, glm_t_2, test = "Chisq")

glm_vpd <- glm(br_status ~ asVPD + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd)

glm_vpd_2 <- glm(br_status ~ poly(asVPD, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_2)

## Square VPD?
anova(glm_vpd, glm_vpd_2, test = "Chisq")


glm_rh <- glm(br_status ~ asRH + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_rh)

glm_rh_2 <- glm(br_status ~ poly(asRH, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_rh_2)

## square rh?
anova(glm_rh, glm_rh_2, test = "Chisq")


glm_p <- glm(br_status ~ asP + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_p)

glm_p_2 <- glm(br_status ~ poly(asP, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_p_2)

## Square P?
anova(glm_p, glm_p_2, test = "Chisq")


glm_t_rh <- glm(br_status ~ asT * asRH + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_rh)

glm_t_2_rh <- glm(br_status ~ poly(asT, 2) * asRH + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2_rh)

anova(glm_t_rh, glm_t_2_rh, test = "Chisq")

glm_t_2_rh_2 <- glm(br_status ~ poly(asT, 2) * poly(asRH, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2_rh_2)

anova(glm_t_2_rh, glm_t_2_rh_2, test = "Chisq")


glm_t_p <- glm(br_status ~ asT * asP + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_p)


glm_t_2_p <- glm(br_status ~ poly(asT, 2) * asP + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2_p)

glm_t_2_p_2 <- glm(br_status ~ poly(asT, 2) * poly(asP, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2_p_2)

anova(glm_t_2_p, glm_t_2_p_2, test = "Chisq")

glm_t_2_p_rh <- glm(br_status ~ poly(asT, 2) * asRH + asP + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2_p_rh)

glm_p_rh <- glm(br_status ~ asP + asRH + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_p_rh)

glm_vpd_p <- glm(br_status ~ asVPD* asP + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_p)

glm_vpd_2_p <- glm(br_status ~ poly(asVPD, 2) * asP + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_2_p)

anova(glm_vpd_p, glm_vpd_2_p, test = "Chisq")

glm_vpd_2_p_2 <- glm(br_status ~ poly(asVPD, 2) * poly(asP, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_2_p_2)

glm_vpd_2_p_2_rh_2 <- glm(br_status ~ poly(asVPD, 2) * poly(asP, 2) * poly(asT, 2) + dbh_cm,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_2_p_2_rh_2)

anova(glm_vpd_2_p, glm_vpd_2_p_2, test = "Chisq")

glm_max_vpd_2 <- glm(br_status ~ poly(max_VPD, 2) + dbh_cm,
    data = mod_data_tree_level,
    family = binomial
)
summary(glm_max_vpd_2)

## model.sel(glm_t, glm_t_2, glm_rh, glm_p, glm_t_rh, glm_t_2_rh, glm_t_p, glm_t_2_p, glm_t_p_rh, glm_t_2_p_rh, glm_p_rh, glm_vpd, glm_vpd_2, glm_vpd_p, glm_vpd_2_p)

## First Order models
model.sel(glm_t,
          glm_rh,
          glm_p,
          glm_t_rh,
          glm_t_p,
          glm_vpd,
          glm_vpd_p)

## Second Order Models
model.sel(glm_t_2,
          glm_rh_2,
          glm_p_2,
          glm_t_2_rh_2,
          glm_t_2_p_2,
          glm_vpd_2,
          glm_vpd_2_p_2)

## All Models
model.sel(glm_t,
          glm_rh,
          glm_p,
          glm_t_rh,
          glm_t_p,
          glm_vpd,
          glm_vpd_p,
          glm_t_2,
          glm_rh_2,
          glm_p_2,
          glm_t_2_rh_2,
          glm_t_p,
          glm_t_2_p,
          glm_vpd_2,
          glm_vpd_2_p,
          glm_vpd_2_p_2)


### No P
## First Order
model.sel(glm_t,
          glm_rh,
          glm_t_rh,
          glm_vpd)

## Second Order
model.sel(glm_t_2,
          glm_rh_2,
          glm_t_2_rh_2,
          glm_vpd_2)

## All
model.sel(glm_t,
          glm_rh,
          glm_t_rh,
          glm_vpd,
          glm_t_2,
          glm_rh_2,
          glm_t_2_rh_2,
          glm_vpd_2)

```

```{r, prediction-plots}
mod_data_tree_level %>%
    select(asT) %>%
    range()

mod_data_tree_level %>%
    select(asRH) %>%
    range()

mod_data_tree_level %>%
    select(asP) %>%
    range()

mod_data_tree_level %>%
    select(asVPD) %>%
    range()

mod_data_tree_level %>%
    select(dbh_cm) %>%
    range()

plot(ggpredict(glm_t_2_p_rh), facet = TRUE)

plot(ggpredict(glm_t_2_p_2), facet = TRUE)

plot(ggpredict(glm_t_2_p_2, terms = c("asP", "asT")), facet = TRUE)

plot(ggpredict(glm_t_2_p_2, terms = c("asP",  "asT [8, 10, 12, 14]")))

plot(ggpredict(glm_vpd_2_p_2), facet = TRUE)

plot(ggpredict(glm_vpd_2_p_2, terms = c("asVPD",  "asP [5, 50, 100, 150, 190]")))

plot(ggpredict(glm_vpd_2_p_2, terms = c("asP", "asVPD [45, 50, 60, 65, 70]")))


plot(ggpredict(glm_t_rh), facet = TRUE)

plot(ggpredict(glm_t_rh, terms = c("asRH", "asT [8, 10, 12, 14]")))


plot(ggpredict(glm_t_2_rh_2), facet = TRUE)

plot(ggpredict(glm_t_2_rh_2, terms = c("asRH", "asT [8, 10, 12, 14]")))


```

### Residual Plots

```{r, residual-plots}
library(car)

glm_vpd_p_park <- glm(br_status ~ asVPD* asP + dbh_cm + park,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_p_park)

glm_vpd_2_p_2_park <- glm(br_status ~ poly(asVPD, 2) * poly(asP, 2) + dbh_cm + park,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_vpd_2_p_2_park)

residualPlots(glm_vpd_p_park, terms = ~ 1 | park)

residualPlots(glm_vpd_2_p_2_park, terms = ~ 1 | park)

glm_t_2_rh_2_park <- glm(br_status ~ poly(asT, 2) * poly(asRH, 2) + dbh_cm + park,
                   data = mod_data_tree_level,
                   family = binomial
                 )
summary(glm_t_2_rh_2_park)

residualPlots(glm_t_2_rh_2_park, terms = ~ 1 | park)



nbc <- 10
cor_r <- pgirmess::correlog(coords=mod_data_tree_level[,c("long", "lat")],
                            z=glm_vpd_2_p_2$residuals,
                            method="Moran", nbclass=nbc)

cor_r

```

### Variogram

```{r variogram}
library(gstat)
library(sp)

library(gstat)

## mod_data_spdf <- SpatialPointsDataFrame(coords = list(lon = mod_data_tree_level$X, lat = mod_data_tree_level$Y), proj4string = toString(crs_lcc), data = mod_data_tree_level)

mod_data_spdf <- as(st_as_sf(mod_data_tree_level, coords = c("X", "Y"), crs = crs_lcc), "Spatial")

## bubble(mod_data_spdf, "asVPD")
## bubble(mod_data_spdf, "asT")
## bubble(mod_data_spdf, "asP")
## bubble(mod_data_spdf, "asRH")


t_semivario <- variogram(asT ~ 1, data = mod_data_tree_level, locations = ~X+Y)
plot(t_semivario)

vpd_semivario <- variogram(asVPD ~ 1, data = mod_data_tree_level, locations = ~X+Y)
plot(vpd_semivario)

rh_semivario <- variogram(asRH ~ 1, data = mod_data_tree_level, locations = ~X+Y)
plot(rh_semivario)

p_semivario <- variogram(asP ~ 1, data = mod_data_tree_level, locations = ~X+Y)
plot(p_semivario)
```

### Confusion Matrix

``` r
library(caret)

set.seed(255)

ctrl <- trainControl(method = "repeatedcv", number = 10, savePredictions = TRUE, repeats = 5)


glm_vpd_2_p_2_train <- train(factor(br_status) ~ poly(asVPD, 2) * poly(asP, 2) + dbh_cm,
    data = mod_data_tree_level,
    method = "glm", family = "binomial",
    trControl = ctrl, tuneLength = 5
    )

glm_t_2_p_2_train <- train(factor(br_status) ~ poly(asT, 2) * poly(asP, 2) + dbh_cm,
    data = mod_data_tree_level,
    method = "glm", family = "binomial",
    trControl = ctrl, tuneLength = 5
    )

glm_p_2_train <- train(factor(br_status) ~ poly(asP, 2) + dbh_cm,
    data = mod_data_tree_level,
    method = "glm", family = "binomial",
    trControl = ctrl, tuneLength = 5
)

glm_t_rh_train <- train(factor(br_status) ~ asT * asRH + dbh_cm,
    data = mod_data_tree_level,
    method = "glm", family = "binomial",
    trControl = ctrl, tuneLength = 5
    )

glm_t_2_rh_2_train <- train(factor(br_status) ~ poly(asT, 2) * poly(asRH, 2) + dbh_cm,
    data = mod_data_tree_level,
    method = "glm", family = "binomial",
    trControl = ctrl, tuneLength = 5
    )

glm_rh_2_train <- train(factor(br_status) ~ poly(asRH, 2) + dbh_cm,
    data = mod_data_tree_level,
    method = "glm", family = "binomial",
    trControl = ctrl, tuneLength = 5
)

print(glm_vpd_2_p_2_train)
print(glm_t_2_p_2_train)
print(glm_p_2_train)
print(glm_t_rh_train)
print(glm_t_2_rh_2_train)
print(glm_rh_2_train)
```

## sdmTMB

``` r
library(sdmTMB)
library(visreg)

## mesh <- make_mesh(mod_data_tree_level, xy_cols = c("X", "Y"), n_knots = 300, type = "kmeans", seed = 255)
## plot(mesh)

## mesh <- make_mesh(mod_data_tree_level,
##                   xy_cols = c("X", "Y"),
##                   fmesher_func = fmesher::fm_mesh_2d_inla,
##                   cutoff = 5,  # minimum triangle edge length
##                   max.edge = c(75, 100), # inner and outer max triangle lengths
##                   offset = c(75,50) # inner and outer border widths
## )
## plot(mesh)

mesh <- make_mesh(mod_data_tree_level, xy_cols = c("X", "Y"), cutoff = 10)
plot(mesh)

## mesh <- make_mesh(mod_data_tree_level,
##                   xy_cols = c("X", "Y"),
##                   fmesher_func = fmesher::fm_mesh_2d_inla,
##                   cutoff = 7.5,  # minimum triangle edge length
##                   max.edge = c(75, 100), # inner and outer max triangle lengths
##                   offset = c(75,100) # inner and outer border widths
## )
## plot(mesh)

sdm_dbh <- sdmTMB(
    br_status ~ log_dbh_cm,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_dbh)

sdm_t <- sdmTMB(
    br_status ~ log_dbh_cm + asT,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t)

sdm_rh <- sdmTMB(
    br_status ~ log_dbh_cm + asRH,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_rh)

sdm_t_rh <- sdmTMB(
    br_status ~ log_dbh_cm + asT + asRH,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_rh)
 
sdm_t_x_rh <- sdmTMB(
    br_status ~ log_dbh_cm + asT * asRH,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_x_rh)

## sdm_t_sumrh <- sdmTMB(
##     br_status ~ log_dbh_cm + asT + sumRH,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_sumrh)

## sdm_t_x_sumrh <- sdmTMB(
##     br_status ~ log_dbh_cm + asT * sumRH,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_x_sumrh)

sdm_t_rhmax <- sdmTMB(
    br_status ~ log_dbh_cm + asT + rollRH_max,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_rhmax)

sdm_t_x_rhmax <- sdmTMB(
    br_status ~ log_dbh_cm + asT * rollRH_max,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_x_rhmax)

sdm_t_p <- sdmTMB(
    br_status ~ log_dbh_cm + asT + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_p)

sdm_t_x_p <- sdmTMB(
    br_status ~ log_dbh_cm + asT * asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_x_p)

sdm_vpdmean <- sdmTMB(
    br_status ~ log_dbh_cm + asVPDmean,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmean)

sdm_vpdmean_p <- sdmTMB(
    br_status ~ log_dbh_cm + asVPDmean + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmean_p)

sdm_vpdmax <- sdmTMB(
    br_status ~ log_dbh_cm + asVPDmax,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmax)

sdm_vpdmax_p <- sdmTMB(
    br_status ~ log_dbh_cm + asVPDmax + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmax_p)

sdm_p <- sdmTMB(
    br_status ~ log_dbh_cm + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_p)

## sdm_t_p_rh <- sdmTMB(
##     br_status ~ log_dbh_cm + asT + asRH + asP,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_p_rh)


## Second Order

sdm_t_2 <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2),
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2)

sdm_t_2_rh <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2) + asRH,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2_rh)

sdm_t_2_x_rh <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2) * asRH,
##    spatial_varying = ~ 0 + asRH,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2_x_rh)

## sdm_t_2_sumrh <- sdmTMB(
##     br_status ~ log_dbh_cm + poly(asT, 2) + sumRH,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_2_sumrh)

## sdm_t_2_x_sumrh <- sdmTMB(
##     br_status ~ log_dbh_cm + poly(asT, 2) * sumRH,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_2_x_sumrh)

sdm_t_2_rhmax <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2) + rollRH_max,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2_rhmax)

sdm_t_2_x_rhmax <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2) * rollRH_max,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2_x_rhmax)


## sdm_t_2_x_rh_1 <- run_extra_optimization(sdm_t_2_x_rh, newton_loops = 100)
## summary(sdm_t_2_x_rh_1)

sdm_vpdmean_2 <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asVPDmean, 2),
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmean_2)

sdm_vpdmean_2_p <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asVPDmean, 2) + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmean_2_p)

sdm_vpdmax_2 <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asVPDmax, 2),
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmax_2)

sdm_vpdmax_2_p <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asVPDmax, 2) + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_vpdmax_2_p)

sdm_t_2_p <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2) + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2_p)

sdm_t_2_x_p <- sdmTMB(
    br_status ~ log_dbh_cm + poly(asT, 2) * asP,
##    spatial_varying = ~ 0 + asP,
    data = mod_data_tree_level,
    family = binomial(link = "logit"),
    mesh = mesh,
    spatial = "on"
)
summary(sdm_t_2_x_p)

## sdm_t_2_p_rh <- sdmTMB(
##     br_status ~ log_dbh_cm + poly(asT, 2) + asRH + asP,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_2_p_rh)

## sdm_t_2_x_rh_p <- sdmTMB(
##     br_status ~ log_dbh_cm + poly(asT, 2) * asRH + asP,
##     data = mod_data_tree_level,
##     family = binomial(link = "logit"),
##     mesh = mesh,
##     spatial = "on"
## )
## summary(sdm_t_2_x_rh_p)


## All models
(model_table <- model.sel(
     sdm_dbh,
     sdm_t,
     sdm_rh,
    sdm_t_rh,
    sdm_t_x_rh,
    ## sdm_t_sumrh,
    ## sdm_t_x_sumrh,
    sdm_t_rhmax,
    sdm_t_x_rhmax,
    sdm_vpdmean,
    sdm_vpdmean_p,
    sdm_vpdmax,
    sdm_vpdmax_p,
    sdm_p,
    sdm_t_2,
    sdm_t_2_rh,
    sdm_t_2_x_rh,
    ## sdm_t_2_sumrh,
    ## sdm_t_2_x_sumrh,
    sdm_t_2_rhmax,
    sdm_t_2_x_rhmax,
    sdm_vpdmean_2,
    sdm_vpdmax_2,
    sdm_vpdmean_2_p,
    sdm_vpdmax_2_p,
    sdm_t_p,
    sdm_t_2_p,
    sdm_t_x_p,
    sdm_t_2_x_p
))


## First Order with interactions
(model_table2 <- model.sel(
     sdm_dbh,
     sdm_t,
     sdm_rh,
    sdm_t_rh,
    sdm_t_x_rh,
    sdm_t_rhmax,
    sdm_t_x_rhmax,
    sdm_vpd,
    sdm_vpd_p,
    sdm_p,
    sdm_t_p,
    sdm_t_x_p
 ))


glm_t_2_x_p <- glm(br_status ~ poly(asT, 2) * asP, data = mod_data_tree_level, family = binomial(link = "logit"))

mod_data_tree_level$resids_glm <- residuals(glm_t_2_x_p)

qqnorm(residuals(glm_t_2_x_p))
qqline(residuals(glm_t_2_x_p))

ggplot(mod_data_tree_level, aes(X, Y, col = resids_glm)) +
  scale_colour_gradient2() +
  geom_point() +
    coord_fixed()
ggsave("resid_map_glm.png")


glmer_t_2_x_p <- lme4::glmer(br_status ~ poly(asT, 2) * asP + (1 | transect_id), data = mod_data_tree_level, family = binomial(link = "logit"))
summary(glmer_t_2_x_p)

mod_data_tree_level$resids_glmer <- residuals(glmer_t_2_x_p)

qqnorm(residuals(glmer_t_2_x_p))
qqline(residuals(glmer_t_2_x_p))

ggplot(mod_data_tree_level, aes(X, Y, col = resids_glmer)) +
  scale_colour_gradient2() +
  geom_point() +
    coord_fixed()
ggsave("resid_map_glmer.png")

mod_data_tree_level$resids <- residuals(sdm_t_2_x_p)

qqnorm(residuals(sdm_t_2_x_p))
qqline(residuals(sdm_t_2_x_p))

ggplot(mod_data_tree_level, aes(X, Y, col = resids)) +
    scale_colour_gradient2() +
    geom_point() +
    ##coord_fixed() +
    facet_wrap(~park, scales = "free")
##ggsave("resid_map_sdmTMB.png")


mod_data_tree_level %>%
    select(asP) %>%
    range()


## visreg(sdm_t_2_x_p, xvar = "asT", by = "asP", gg = TRUE, overlay = TRUE, scale = "response")

## visreg(sdm_t_x_rh, xvar = "asRH", by = "asT", gg = TRUE, overlay = TRUE) ## Log Scale

## visreg(sdm_t_x_rh, xvar = "asRH", by = "asT", gg = TRUE, overlay = TRUE, scale = "response")

## visreg(sdm_t_2_x_rh, xvar = "asRH", by = "asT", gg = TRUE, overlay = TRUE, scale = "response")

plot(ggeffect(sdm_t), facet = TRUE, add.data = TRUE)

plot(ggeffect(sdm_t_x_rh), facet = TRUE, add.data = TRUE)

plot(ggeffect(sdm_t_x_rh, terms = c("asT", "asRH [30, 40, 50, 60]")), add.data = TRUE)

plot(ggeffect(sdm_t_x_rh, terms = c("asRH", "asT [8:14 by=2]")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_rh), facet = TRUE, add.data = TRUE)

plot(ggeffect(sdm_t_2_x_rh, terms = c("log_dbh_cm")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_rh, terms = c("asRH")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_rh, terms = c("asT")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_rh, terms = c("asT", "asRH [30,40,50,60]")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_rh, terms = c("asRH", "asT [8:14 by=1]")), add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p), facet = TRUE, add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p, terms = c("dbh_cm")), add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p, terms = c("asT")), add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p, terms = c("asRH")), add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p, terms = c("asP")), add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p, terms = c("asT", "asRH [30,40,50,60]")), add.data = TRUE)

## plot(ggeffect(sdm_t_2_x_rh_p, terms = c("asRH", "asT [8:14 by=1]")), add.data = TRUE)

plot(ggeffect(sdm_t_x_p), facet = TRUE, add.data = TRUE)

plot(ggeffect(sdm_t_x_p, terms = c("asT", "asP[25:175 by = 50]")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_p), facet = TRUE, add.data = TRUE)

plot(ggeffect(sdm_t_2_x_p, terms = c("asT", "asP[25:175 by = 50]")), add.data = TRUE)

plot(ggeffect(sdm_t_2_x_p, terms = c("asP", "asT [8:14 by=1]")), add.data = TRUE)


plot(ggeffect(sdm_t_rhmax), facet = TRUE, add.data = TRUE)

plot(ggeffect(sdm_t_rhmax, terms = c("asT", "rollRH_max")), add.data = TRUE)


sdmTMB_cv(formula(sdm_t_2_x_p),
          data = mod_data_tree_level,
          mesh = mesh,
          family = binomial(link = "logit")
          )

sdmTMB_cv(formula(sdm_t_2_x_rh),
          data = mod_data_tree_level,
          mesh = mesh,
          family = binomial(link = "logit")
          )

tidy(sdm_t_x_p, conf.int = TRUE)

tidy(sdm_t_2_x_p, conf.int = TRUE)

tidy(sdm_t_x_rh, conf.int = TRUE)

tidy(sdm_t_2_x_rh, conf.int = TRUE)

sim_dat <- simulate(
    sdm_t_2_x_p
)

predictor_dat <- expand.grid(
    X = seq(-419, 657, length.out = 400), Y = seq(4431, 5818, length.out = 400)
)

p <- predict(sdm_t_2_x_p, newdata = predictor_dat, nsim = 500)


saveRDS(sdm_t_2_x_p, file = "sdm_t_2_x_p.RDS")

```

## Conf Matrix

``` r
library(cvms)

set.seed(255)

clust <- sample(1:8, size = nrow(mod_data_tree_level), replace = T)

sdm_cv_t_2_x_p <- sdmTMB_cv(formula(sdm_t_2_x_p),
    data = mod_data_tree_level,
    mesh = mesh,
    fold_ids = clust,
    family = binomial(link = "logit"),
    k_folds = length(unique(clust))
)

prediction_t_2_x_p <- sdm_cv_t_2_x_p$data %>%
    mutate(cv_predicted_class = if_else(cv_predicted >= 0.5, 1, 0))

confusion_matrix(
    targets = prediction_t_2_x_p$br_status,
    predictions = prediction_t_2_x_p$cv_predicted_class
) %>%
    plot_confusion_matrix(font_counts = font(size = 5),
                          font_normalized = font(size = 5),
                          font_row_percentages = font(size = 5),
                          font_col_percentages = font(size = 5)) +
    theme(text = element_text(size = 20))


sdm_cv_t_2_x_rh <- sdmTMB_cv(formula(sdm_t_2_x_rh),
    data = mod_data_tree_level,
    mesh = mesh,
    fold_ids = clust,
    family = binomial(link = "logit"),
    k_folds = length(unique(clust))
)

prediction_t_2_x_rh <- sdm_cv_t_2_x_rh$data %>%
    mutate(cv_predicted_class = if_else(cv_predicted >= 0.5, 1, 0))

confusion_matrix(
    targets = prediction_t_2_x_rh$br_status,
    predictions = prediction_t_2_x_rh$cv_predicted_class
) %>%
    plot_confusion_matrix(font_counts = font(size = 5),
                          font_normalized = font(size = 5),
                          font_row_percentages = font(size = 5),
                          font_col_percentages = font(size = 5)) +
    theme(text = element_text(size = 20))

sdm_cv_t <- sdmTMB_cv(formula(sdm_t),
    data = mod_data_tree_level,
    mesh = mesh,
    fold_ids = clust,
    family = binomial(link = "logit"),
    k_folds = length(unique(clust))
)

prediction_t <- sdm_cv_t$data %>%
    mutate(cv_predicted_class = if_else(cv_predicted >= 0.5, 1, 0))

confusion_matrix(
    targets = prediction_t$br_status,
    predictions = prediction_t$cv_predicted_class
) %>%
    plot_confusion_matrix(
        font_counts = font(size = 5),
        font_normalized = font(size = 5),
        font_row_percentages = font(size = 5),
        font_col_percentages = font(size = 5)
    ) +
    theme(text = element_text(size = 20))

sdm_cv_t_2_x_p$sum_loglik
sdm_cv_t_2_x_rh$sum_loglik
sdm_cv_t$sum_loglik

sdmTMB_stacking(list(sdm_cv_t_2_x_p, sdm_cv_t_2_x_rh))
```

